\section{Conclusion}
\label{sec:conclusion}

We presented the first direct measurement of steering vector decay dynamics in
language models. Our experiments on \gptxl across three behavioral dimensions
reveal that representational decay is instantaneous: within a single generation
step after steering removal, hidden states at the injection layer are
indistinguishable from those of an unsteered model. This finding is universal
across all layers, multiplier strengths, and steering durations tested.

The key insight is that steering influence decomposes into two components with
vastly different timescales. Representational influence (hidden state alignment
with the steering direction) has a half-life of less than one token. Behavioral
influence (divergence of output distributions from the unsteered baseline)
persists for over 75~tokens, but operates entirely through the autoregressive
token feedback loop---not through hidden state dynamics.

These findings have direct practical implications: continuous vector application
is necessary for sustained representational steering, and the generated token
sequence is the primary carrier of long-term behavioral influence. For
theorists, the instantaneous absorption of perturbations reveals that
transformer hidden states are remarkably robust, with normalization and residual
connections projecting perturbed states back onto the natural activation
manifold within a single forward pass.

Future work should test whether these dynamics hold for larger models, whether
multi-layer steering creates more persistent effects, and whether alternative
injection targets (\eg key-value caches) exhibit different decay properties.
Understanding decay across architectures and scales will be essential for
designing efficient and reliable steering methods.
